#Predictors of National Carbon Dioxide Emissions
##by Nat Goodby & James Valadez

Are urbanization rates and GDP predictors of a country's carbon dioxide emmissions?

The data comes from the [CIA World Factbook](https://www.cia.gov/library/publications/the-world-factbook/)
Data was cleaned in Python. That process can be seen [here](https://github.com/ngoodby/Final_Project_James_Nat/blob/master/Data/Data_Preprocessing.ipynb)

Let's call the data:
```{r, eval=TRUE}
dataLink="https://raw.githubusercontent.com/ngoodby/Final_Project_James_Nat/master/Data/Cleaned_Data.csv"
df=read.csv(dataLink,stringsAsFactors = FALSE)
```
And look at the contents:
```{r, eval=TRUE}
names(df)
```
```{r}
str(df)
```
What countries have the highest CO2 emissions? 
```{r}
head(df[order(df$CO2_Emissions_PerCapita, decreasing= T),], 10)
```
Let's get a sense of how far outlying Gibraltar's CO2 Emissions are:
```{r}
plot(df$Urbanization_Percentage,df$CO2_Emissions_PerCapita)
```
Would be more helpful to see this plot without Gibraltar. Let's locate it, drop it, and save the resulting dataframe: 
```{r}
df[76,]
```
```{r}
df_noGib=df[-76, ]
```
Now what does that look like if we plot again:
```{r}
plot(df_noGib$Urbanization_Percentage,df_noGib$CO2_Emissions_PerCapita)
```
```{r}
plot(df_noGib$GDP_Per_Capita_Number,df_noGib$CO2_Emissions_PerCapita)
```
Looks like there are higher CO2 emissions per capita as Urbanization rates and GDP per capita increase but let's look at the correlation strength:
```{r}
cor(df_noGib$Urbanization_Percentage, df_noGib$CO2_Emissions_PerCapita)
```
```{r}
cor(df_noGib$GDP_Per_Capita_Number, df_noGib$CO2_Emissions_PerCapita)
```
GDP_Per_Capita is showing higher correlation with CO2 Emissions.
```{r}
test1=lm(as.numeric(Urbanization_Percentage)~CO2_Emissions_PerCapita,data=df_noGib)
summary(test1)
```
Those two variables show *significance* however the adjusted R-squared value is only around 0.25.
```{r}
test2=lm(as.numeric(GDP_Per_Capita_Number)~CO2_Emissions_PerCapita,data=df_noGib)
summary(test2)
```
GDP per capita shows a higher correlation (adjusted R-squared of 0.54) with CO2 emissions than Urbanization percentage did.
Looking at how both independent varibles affect CO2 emissions collectively:
```{r}
test3=lm(CO2_Emissions_PerCapita ~ GDP_Per_Capita_Number + Urbanization_Percentage, data=df_noGib)
summary(test3)
```

